Quantum Pseudo-Telepathy

### 2023-06-29 paper/main.tex
circuits row

\newcommand{\ketbra}[2]{\ensuremath{| #1 \rangle \langle #2 |}}
\newcommand{\ci}{\mathrm{i}}
\DeclarePairedDelimiter{\card}{\lvert}{\rvert}
\DeclarePairedDelimiter{\norm}{\lVert}{\rVert}
\DeclareMathOperator{\tr}{tr}

@@

\begin{figure}[htbp]
  \centering
  \begin{quantikz}
    \lstick{\ket{0}} & \gate{H}
    \slice[label style={xshift=-15mm}]
    {\(\frac{1}{2}(\ket{00} + \ket{01} + \ket{10} +
      \ket{11}) \otimes \ket{00}\)}
    &&&&& \ctrl{2} &&
    \\
    \lstick{\ket{0}} & \gate{H} &&&&&& \ctrl{2} & \\
    \lstick{\ket{0}} & &&&&& \targ{} && \\
    \lstick{\ket{0}} & &&&&&& \targ{}
    \slice[label style={xshift=5mm}]
    {\(\frac{1}{2}(\ket{0000} + \ket{0101} + \ket{1010} +
      \ket{1111})\)}
    &
  \end{quantikz}
  \caption{\label{fig:circuit-psi}}
\end{figure}

As seen in Section~\ref{sec:classic-sol}, the \emph{magic} square is
called this way because, under classical assumptions, there is no such
a square: it is impossible to construct a \(3 \times 3\) matrix filled
with \(\pm 1\)'s, with the constraints that the product of every row
is \(+1\) and the product of every column is \(-1\). We can, however,
construct a ``quantum magic'' square. Figure~\ref{fig:quantum} shows
the \(3 \times 3\) grid that Alice and Bob can use to win the Magic
Square game systematically.  Notice that the eigenvalues of the
observables at each cell are \(\pm 1\). Furthermore, the product of
every row is \(+I\) and the product of every column is \(-I\).  The
quantum strategy works as follows: Alice measures all the observables
from the \(i\)-th row of the predefined matrix. She fills each cell
with the result of each measurement. Similarly, Bob measures all the
observables in the \(j\)-th column that was assigned to him, and fills
each cell of his column with the outcomes of each
measurement. Example~\ref{ex:quantum-sol} depicts a round of the game.

{\renewcommand{\arraystretch}{2}
\begin{figure}
  \centering
  \[\begin{array}{|*{3}{>{\quad}c<{\quad}|}>{\quad}l<{\quad}}
      \hhline{---~}
      +I \otimes Z & +Z \otimes I & +Z \otimes Z & \prod = I \otimes I = I
      \\ \hhline{---~}
      +X \otimes I & +I \otimes X & +X \otimes X & \prod = I \otimes I = I
      \\ \hhline{---~}
      -X \otimes Z & -Z \otimes X & +Y \otimes Y & \prod = \ci{}I \otimes -\ci{}I = I
      \\ \hhline{---~}
      \multicolumn{1}{>{\quad}c<{\quad}}{\prod = -I \otimes I = -I}
      & \multicolumn{1}{>{\quad}c<{\quad}}{\prod = -I \otimes I = -I}
      & \multicolumn{1}{>{\quad}c<{\quad}}{\prod = -\ci{}I \otimes -\ci{}I = -I}
    \end{array}\]
  \caption{\label{fig:quantum}}
\end{figure}}

@@

\subsection{Formal Proof}
To prove that the quantum solution indeed works, we must demonstrate
two fundamental properties. First, the product of the outcomes of the
three measurements on each row (equivalently, column) is equal to 1
(equiv. -1). Second, the outcome of Alice's measurements on the
intersection is always equal to the outcome of Bob's
measurement. Lemma~\ref{lem:product} formalizes the product property.

@@

\begin{proof}
  To prove that Alice and Bob agree at the intersection, it suffices
  to look to the eigenvectors shared by each set of mutually commuting
  operators, shown in Tables~\ref{tab:eigenv-rows} and
  \ref{tab:eigenv-cols}.  Notice that, regardless of the row assigned
  to Alice, the state of her subsystem after her measurements is
  always one of the eigenvectors in that row. This state is always a
  linear combination of one of Bob's eigenspaces for the cell in the
  intersection (the same applies if we consider first Bob than Alice).
  Take, for instance, the second row and first column. Suppose that
  Alice's measurements collapsed her qubits to
  \((\nicefrac{1}{2}) (\ket{00} - \ket{01} - \ket{10} + \ket{11})\).
  This state can be written as
  \((\nicefrac{1}{\sqrt{2}}) (\ket{00} - \ket{10}) -
  (\nicefrac{1}{\sqrt{2}}) (\ket{01} - \ket{11})\), a linear
  combination of the \(-1\)-eigenspace depicted in
  Table~\ref{tab:eigenv-cols} (column 1, row 2).  What is left is to
  show that Bob's state will be exactly the same as Alice's, implying
  that Bob shall observe the same result. This follows from the fact
  that the initial state is already symmetric (the first qubit of
  Alice is equal to the first qubit of Bob; the same is true for their
  second qubits) and the symmetry is preserved by any measurement
  performed by Alice and Bob. For the first row, it is easy to see.
  The others are trickier, but it all boils down to doing the
  calculations.  Consider, for instance, the second row, and suppose
  that the outcomes were \(-1\), \(-1\) and \(+1\), so that Alice's
  state collapsed to
  \((\nicefrac{1}{2}) (\ket{00} - \ket{01} - \ket{10} +
  \ket{11})\). The state after the first measurement is
  \begin{align}
    &\frac{1}{2\sqrt{2}} \Bigl(
    (\ket{00} - \ket{10}) \ket{00} +
    (\ket{01} - \ket{11}) \ket{01} +
    (\ket{10} - \ket{00}) \ket{10} +
    (\ket{11} - \ket{01}) \ket{11}
    \Bigr) \nonumber \\
    \equiv\;
    & \frac{1}{2\sqrt{2}} \Bigl(
      \ket{00} (\ket{00} - \ket{10}) +
      \ket{01} (\ket{01} - \ket{11}) +
      \ket{10} (\ket{10} - \ket{00}) +
      \ket{11} (\ket{11} - \ket{01})
    \Bigr).
  \end{align}
  And, the state after the three measurements is
  \begin{alignat}{2}
    &\frac{1}{4} \Bigl(
      &&(\ket{00} - \ket{01} - \ket{10} + \ket{11}) \ket{00} -
        (\ket{00} - \ket{01} - \ket{10} + \ket{11}) \ket{01} - \nonumber \\
    & &&(\ket{00} - \ket{01} - \ket{10} + \ket{11}) \ket{10} +
      (\ket{00} - \ket{01} - \ket{10} + \ket{11}) \ket{11}
    \Bigr) \nonumber \\
    \equiv\;
    & \frac{1}{4} \Bigl(
      &&\ket{00} (\ket{00} - \ket{01} - \ket{10} + \ket{11}) -
        \ket{01} (\ket{00} - \ket{01} - \ket{10} + \ket{11}) - \nonumber \\
    & &&\ket{10} (\ket{00} - \ket{01} - \ket{10} + \ket{11}) +
        \ket{11} (\ket{00} - \ket{01} - \ket{10} + \ket{11})
    \Bigr).
  \end{alignat}
  The calculations for the other rows and columns are similar, so we omitted them.
  \qed
\end{proof}
\section{Quantum Unitaries (? research)}
\label{sec:quantum-unitaries}

@@

>>>>

\newcommand{\ketbra}[2]{\ensuremath{| #1 \rangle \langle #2 |}}
\newcommand{\ci}{\mathrm{i}}
\DeclarePairedDelimiter{\card}{\lvert}{\rvert}
\DeclarePairedDelimiter{\norm}{\lVert}{\rVert}
\DeclareMathOperator{\tr}{tr}
\newcommand{\cnot}{\mathit{CNOT}}

@@

\begin{figure}[htbp]
  \centering
  \begin{quantikz}
    \lstick{\ket{0}} & \gate{H}
    \slice[label style={xshift=-15mm}]
    {\(\frac{1}{2}(\ket{00} + \ket{01} + \ket{10} +
      \ket{11}) \otimes \ket{00}\)}
    &&&&& \ctrl{2} &&
    \\
    \lstick{\ket{0}} & \gate{H} &&&&&& \ctrl{2} & \\
    \lstick{\ket{0}} & &&&&& \targ{} && \\
    \lstick{\ket{0}} & &&&&&& \targ{}
    \slice[label style={xshift=5mm}]
    {\(\frac{1}{2}(\ket{0000} + \ket{0101} + \ket{1010} +
      \ket{1111})\)}
    &
  \end{quantikz}
  \caption{Quantum circuit for preparing the initial joint state
    of Alice and Bob.\label{fig:circuit-psi}}
\end{figure}

As seen in Section~\ref{sec:classic-sol}, the \emph{magic} square is
called this way because, under classical assumptions, there is no such
a square: it is impossible to construct a \(3 \times 3\) matrix filled
with \(\pm 1\)'s, with the constraints that the product of every row
is \(+1\) and the product of every column is \(-1\). We can, however,
construct a ``quantum magic'' square. Figure~\ref{fig:quantum} shows
the \(3 \times 3\) grid that Alice and Bob can use to win the Magic
Square game systematically.  Notice that the eigenvalues of the
observables at each cell are \(\pm 1\). Furthermore, the product of
every row is \(+I\) and the product of every column is \(-I\).  The
quantum strategy works as follows: Alice measures all the observables
from the \(i\)-th row of the predefined matrix. She fills each cell
with the result of each measurement. Similarly, Bob measures all the
observables in the \(j\)-th column that was assigned to him, and fills
each cell of his cells with the outcomes of each
measurement. Example~\ref{ex:quantum-sol} depicts a round of the game.

{\renewcommand{\arraystretch}{2}
\begin{figure}
  \centering
  \[\begin{array}{|*{3}{>{\quad}c<{\quad}|}>{\quad}l<{\quad}}
      \hhline{---~}
      +I \otimes Z & +Z \otimes I & +Z \otimes Z & \prod = I \otimes I = I
      \\ \hhline{---~}
      +X \otimes I & +I \otimes X & +X \otimes X & \prod = I \otimes I = I
      \\ \hhline{---~}
      -X \otimes Z & -Z \otimes X & +Y \otimes Y & \prod = \ci{}I \otimes -\ci{}I = I
      \\ \hhline{---~}
      \multicolumn{1}{>{\quad}c<{\quad}}{\prod = -I \otimes I = -I}
      & \multicolumn{1}{>{\quad}c<{\quad}}{\prod = -I \otimes I = -I}
      & \multicolumn{1}{>{\quad}c<{\quad}}{\prod = -\ci{}I \otimes -\ci{}I = -I}
    \end{array}\]
  \caption{Observables used by Alice and Bob to win the Magic Square game.\label{fig:quantum}}
\end{figure}}

@@

\subsection{Formal Proof}
To prove that the quantum solution indeed works, we must demonstrate
two fundamental properties. First, the product of the outcomes of the
three measurements on each row (equivalently, column) is equal to 1
(equiv. -1). Second, the outcome of Alice's measurement on the
intersection is always equal to the outcome of Bob's
measurement. Lemma~\ref{lem:product} formalizes the product property.

@@

\begin{proof}
  To prove that Alice and Bob agree at the intersection, it suffices
  to look to the eigenvectors shared by each set of mutually commuting
  operators, shown in Tables~\ref{tab:eigenv-rows} and
  \ref{tab:eigenv-cols}.  Notice that, regardless of the row assigned
  to Alice, the state of her subsystem after her measurements is
  always one of the eigenvectors in that row. This state is always a
  linear combination of one of Bob's eigenspaces for the cell in the
  intersection (the same applies if we consider first Bob than Alice).
  Take, for instance, the second row and first column. Suppose that
  Alice's measurements collapsed her qubits to
  \((\nicefrac{1}{2}) (\ket{00} - \ket{01} - \ket{10} + \ket{11})\).
  This state can be written as
  \((\nicefrac{1}{\sqrt{2}}) (\ket{00} - \ket{10}) -
  (\nicefrac{1}{\sqrt{2}}) (\ket{01} - \ket{11})\), a linear
  combination of the \(-1\)-eigenspace depicted in
  Table~\ref{tab:eigenv-cols} (column 1, row 2).  What is left is to
  show that Bob's state will be exactly the same as Alice's, implying
  that Bob shall observe the same result. This follows from the fact
  that the initial state is already symmetric (the first qubit of
  Alice is equal to the first qubit of Bob; the same is true for their
  second qubits) and the symmetry is preserved by any measurement
  performed by Alice and Bob. For the first row, it is easy to see.
  The others are trickier, but it all boils down to doing the
  calculations.  Consider, for instance, the second row, and suppose
  that the outcomes were \(-1\), \(-1\) and \(+1\), so that Alice's
  state collapsed to
  \((\nicefrac{1}{2}) (\ket{00} - \ket{01} - \ket{10} +
  \ket{11})\). The state after the first measurement is
  \begin{align}
    &\frac{1}{2\sqrt{2}} \Bigl(
    (\ket{00} - \ket{10}) \ket{00} +
    (\ket{01} - \ket{11}) \ket{01} +
    (\ket{10} - \ket{00}) \ket{10} +
    (\ket{11} - \ket{01}) \ket{11}
    \Bigr) \nonumber \\
    \equiv\;
    & \frac{1}{2\sqrt{2}} \Bigl(
      \ket{00} (\ket{00} - \ket{10}) +
      \ket{01} (\ket{01} - \ket{11}) +
      \ket{10} (\ket{10} - \ket{00}) +
      \ket{11} (\ket{11} - \ket{01})
    \Bigr).
  \end{align}
  And, the state after the second measurement is
  \begin{alignat}{2}
    &\frac{1}{4} \Bigl(
      &&(\ket{00} - \ket{01} - \ket{10} + \ket{11}) \ket{00} -
        (\ket{00} - \ket{01} - \ket{10} + \ket{11}) \ket{01} - \nonumber \\
    & &&(\ket{00} - \ket{01} - \ket{10} + \ket{11}) \ket{10} +
      (\ket{00} - \ket{01} - \ket{10} + \ket{11}) \ket{11}
    \Bigr) \nonumber \\
    \equiv\;
    & \frac{1}{4} \Bigl(
      &&\ket{00} (\ket{00} - \ket{01} - \ket{10} + \ket{11}) -
        \ket{01} (\ket{00} - \ket{01} - \ket{10} + \ket{11}) - \nonumber \\
    & &&\ket{10} (\ket{00} - \ket{01} - \ket{10} + \ket{11}) +
        \ket{11} (\ket{00} - \ket{01} - \ket{10} + \ket{11})
    \Bigr).
  \end{alignat}
  The calculations for the other rows and columns are similar, so we omitted them.
  \qed
\end{proof}

\subsection{Implementing with Circuits}
In this section, we show the circuits used to implement the quantum
strategy. We implemented the solution using the \texttt{Qiskit}
library. \texttt{Qiskit} only allows measurements in the computational
basis. For any other basis \(B\), we must find an unitary
transformation \(U\) such that \(B = U^{*} (Z \otimes I) U\) (where
\(U^{*}\) denotes the Hermitian adjoint of \(U\)). For example,
\(Z \otimes Z = \cnot{} (I \otimes Z) \cnot{}\) and
\(X \otimes X = (H \otimes H) (Z \otimes Z) (H \otimes H)\). Figure~\ref{fig:change-basis}
shows the unitaries for each of the nine observables.

{\renewcommand{\arraystretch}{2}
\begin{figure}
  \centering
  \[\begin{array}{|*{2}{>{\quad}c<{\quad}|}>{\quad}l<{\quad}|}
      \hhline{---}
      (I \otimes I) (Z \otimes I) (I \otimes I) & (I \otimes I) (I \otimes Z) (I \otimes I) & \cnot (I \otimes Z) \cnot
      \\ \hhline{---}
      (H \otimes I) (Z \otimes I) (H \otimes I) & (I \otimes H) (I \otimes Z) (I \otimes H) & (H \otimes H) (Z \otimes Z) (H \otimes H)
      \\ \hhline{---}
      (H \otimes Y) (Z \otimes Z) (H \otimes Y) & (Y \otimes H) (Z \otimes Z) (Y \otimes H) &
       (S \otimes S) (X \otimes X) (S^{*} \otimes S^{*}) 
      \\ \hhline{---}
    \end{array}\]
  \caption{Unitary transformations required for measuring in the
    different basis. The actual basis measured is the middle
    term of each cell, which, after expanding, is either
    \(Z \otimes I\) or \(I \otimes Z\). For example, in the last
    cell, \(X \otimes X\) evaluates to
    \((H \otimes H) (Z \otimes Z) (H \otimes H)\).
    \(Z \otimes Z\), in turn, becomes \(\cnot (I \otimes Z) \cnot\), meaning that
    we measure the second qubit in the computational basis.\label{fig:change-basis}}
\end{figure}}

We now show the quantum circuits for each row and each column:

\begin{description}
\item[Row 0:] When assigned the first row, Alice performs the measurements
  \(I \otimes Z, Z \otimes I\) and \(Z \otimes Z\), which the circuit
  in Figure~\ref{fig:circ-row0} implements.

  \begin{figure}[htbp]
    \centering
    \begin{quantikz}
      \lstick{\(A_{0}\)}
      & \metercw[label style={inner sep=1pt}]{a_{1}}
      & \targ{}
      &
      & \targ{}
      &
      \\
      \lstick{\(A_{1}\)}
      & \metercw[label style={inner sep=1pt}]{a_{0}}
      & \ctrl{-1}
      & \metercw[label style={inner sep=1pt}]{a_{2}}
      & \ctrl{-1}
      &
    \end{quantikz}
    \caption{Quantum circuit for measuring the first row.\label{fig:circ-row0}}
  \end{figure}

  \item[Row 1:] When assigned the second row, Alice performs the measurements
  \(X \otimes I, I \otimes X\) and \(X \otimes X\), which the circuit
  in Figure~\ref{fig:circ-row1} implements.

  \begin{figure}[htbp]
    \centering
    \begin{quantikz}
      \lstick{\(A_{0}\)}
      & \gate{H}
      & \metercw[label style={inner sep=1pt}]{a_{0}}
      & \ctrl{1}
      &
      & \ctrl{1}
      & \gate{H}
      &
      \\
      \lstick{\(A_{1}\)}
      & \gate{H}
      & \metercw[label style={inner sep=1pt}]{a_{1}}
      & \targ{}
      &  \metercw[label style={inner sep=1pt}]{a_{2}}
      & \targ{}
      & \gate{H}
      &
    \end{quantikz}
    \caption{Quantum circuit for measuring the second row.\label{fig:circ-row1}}
  \end{figure}

    \item[Row 2:] When assigned the third row, Alice performs the measurements
  \(-X \otimes Z, -Z \otimes X\) and \(Y \otimes Y\), which the circuit
  in Figure~\ref{fig:circ-row2} implements.

  \begin{figure}[htbp]
    \centering
    \begin{quantikz}[column sep=0.25cm]
      \lstick{\(A_{0}\)}
      & \gate{H}
      & \ctrl{1}
      &
      & \ctrl{1}
      & \gate{H}
      & \gate{Y}
      & \ctrl{1}
      &
      & \ctrl{1}
      & \gate{S}
      & \gate{H}
      & \ctrl{1}
      &
      & \ctrl{1}
      & \gate{H}
      & \gate{S^*}
      \\
      \lstick{\(A_{1}\)}
      & \gate{Y}
      & \targ{}
      & \metercw[label style={inner sep=1pt}]{a_{0}}
      & \targ{}
      & \gate{Y}
      & \gate{H}
      & \targ{}
      & \metercw[label style={inner sep=1pt}]{a_{1}}
      & \targ{}
      & \gate{S}
      & \gate{H}
      & \targ{}
      & \metercw[label style={inner sep=1pt}]{a_{2}}
      & \targ{}
      & \gate{H}
      & \gate{S^*}
    \end{quantikz}
    \caption{Quantum circuit for measuring the third row.\label{fig:circ-row2}}
  \end{figure}
\end{description}

@@

\section{Quantum Unitaries (? research)}
\label{sec:quantum-unitaries}
